---
title: "Final Project"
format: html
editor: visual
bibliography: references.bib
link-citations: true
---

Final Project

1.  Load Packages

```{r}
library(Lahman)
library(nlme)
library(janitor)
library(tidyverse)
library(rio)
library(mdsr)
library(here)
library(ggplot2)
```

2.  Load Data

```{r}
data <- import(here("data", "finalproj.csv")) %>%
  as_tibble()
```

Data Cleaning- Demographics/Characteristics

```{r}
data$gender <- as.factor(data$gender)
data$gender

data$age <- as.factor(data$age)
data$age

data <- data%>%
  mutate(id= row_number())

data <- data %>%
  mutate(yrs_combine = recode(
    yrs_teach,
    "0-1 years" = "0-5 years",
    "2-5 years" = "0-5 years",
    "6-10 years" = "6-10 years",
    "11-15 years" = "11+ years",
    "More than 15 years" = "11+ years",
    .default ="Unknown"  # Change as necessary for unmatched values
  ))

data
```

Measure Cleaning

```{r}
erq_data_recode <- data %>%
  mutate(across(starts_with("er"), ~ case_when(.== "Strongly disagree" ~ 1,
                                               .=="Disagree" ~ 2,
                                               .=="Somewhat disagree" ~ 3,
                                               .=="Neither agree nor disagree (neutral)" ~ 4,
                                               .=="Somewhat agree" ~ 5,
                                               .=="Agree" ~ 6,
                                               .=="Strongly agree" ~ 7,
                                               .== "" ~ NA_real_)))

pro_data_recode <- erq_data_recode %>%
  mutate(across(starts_with("pro"), ~ case_when(.=="Never" ~ 1,
                                               .=="Rarely" ~ 2,
                                               .=="Sometimes" ~ 3,
                                               .=="Often" ~ 4,
                                               .=="Very often" ~ 5,
                                               .== "" ~ NA_real_)))

final_data <- pro_data_recode %>%
  mutate(across(starts_with("mhl"), ~ case_when(.== "Strongly disagree" ~ 1,
                                               .=="Disagree" ~ 2,
                                               .=="Somewhat disagree" ~ 3,
                                               .=="Somewhat agree" ~ 4,
                                               .=="Agree" ~ 5,
                                               .=="Strongly agree" ~ 6,
                                               .== "" ~ NA_real_)))



final_data<- final_data %>%
  mutate(mean_mhl_id = ((mhl_1 + mhl_2 + mhl_3 + mhl_4 + mhl_5 + mhl_6 + mhl_7 + mhl_8 + mhl_9 + mhl_10 + mhl_11 + mhl_11 + mhl_12 + mhl_13 + mhl_14 + mhl_15)/15))

final_data
```

ProQOL

```{r}
#Step 1: Reversals

#items 1, 2, 15, 17, 29
# 1 ->5
# 2 ->4
# 4-> 2
# 5 ->1

final_data <- final_data %>%
  mutate(pro_1 = as.numeric(recode(
    pro_1,
    `1` = "5",
    `2` = "4",
    `3` = "3",
    `4` = "2",
    `5` = "1")))

final_data <- final_data %>%
  mutate(pro_2 = as.numeric(recode(
    pro_2,
    `1` = "5",
    `2` = "4",
    `3` = "3",
    `4` = "2",
    `5` = "1")))

final_data <- final_data %>%
  mutate(pro_15 = as.numeric(recode(
    pro_15,
    `1` = "5",
    `2` = "4",
    `3` = "3",
    `4` = "2",
    `5` = "1")))

final_data <- final_data %>%
  mutate(pro_17 = as.numeric(recode(
    pro_17,
    `1` = "5",
    `2` = "4",
    `3` = "3",
    `4` = "2",
    `5` = "1")))

final_data <- final_data %>%
  mutate(pro_29 = as.numeric(recode(
    pro_29,
    `1` = "5",
    `2` = "4",
    `3` = "3",
    `4` = "2",
    `5` = "1")))

#Step 2 Compassion Satisfaction

#add scores on 3, 6, 12, 16, 18, 20, 22, 24, 27, 30

final_data <- final_data %>%
  mutate(sum_comp_satis = (pro_3 + pro_6 + pro_12 + pro_16 + pro_18 + pro_20 + pro_22 + pro_24 + pro_27 + pro_30))

# Step 3 Burnout 
#add scores on 1, 4, 8, 10, 15, 17, 19, 21, 26, 29 
final_data

final_data <- final_data %>%
  mutate(sum_burnout = pro_1 + pro_4 + pro_8 + pro_10 + pro_15 + pro_17 + pro_19 + pro_21 + pro_26 + pro_29)


# Step 4 Secondary Traumatic Stress
#add scores on 2, 5, 7, 9, 11, 13, 14, 23, 25, 28 

final_data <- final_data %>%
  mutate(sum_sectrauma = pro_2 + pro_5 + pro_7 + pro_9 + pro_11 + pro_13 + pro_14 + pro_23 + pro_25 + pro_28)

#step 5 frequency table and histogram for compassion

frequency_table <- final_data %>%
  count(sum_comp_satis) %>%
  arrange(desc(n))
frequency_table
print(frequency_table)


comp_hist <- ggplot(final_data, aes(x=sum_comp_satis)) +
  geom_histogram()

comp_hist

#step 6 frequency table and histogram for burnout

frequency_table_2 <- final_data %>%
  count(sum_burnout) %>%
  arrange(desc(n))
frequency_table_2
print(frequency_table_2)


burnout_hist <- ggplot(final_data, aes(x=sum_burnout)) +
  geom_histogram()

burnout_hist

#step 7 frequency table and histogram for trauma 

frequency_table_3 <- final_data %>%
  count(sum_sectrauma) %>%
  arrange(desc(n))
frequency_table_3
print(frequency_table_3)


trauma_hist <- ggplot(final_data, aes(x=sum_sectrauma)) +
  geom_histogram()

trauma_hist
```

Research Questions

1.  Do teacher reports of mental health efficacy differ by school type (i.e., elementary, middle, high school)? 

2.  Is there a relationship between teacher years of experience and teacher reports of mental health efficacy?

3.  Is there a relationship between teacher mental health efficacy and burnout?

Test
